Generative Dynamics of Supreme Court Citations: Analysis with a New Statistical Network Model

### 2020-08-18 Tex/PA_rr/main.tex
expanded on tempral component of hub effect

We first discuss the dependence effect estimates in Figure \ref{SCC_results_1}. In most of the years that it is estimable, the reciprocity effect is statistically significantly positive, and substantial in magnitude. In a typical year, the presence of a citation from case $i$ to case $j$ increases the log-odds of a citation from case $j$ to case $i$ by approximately 0.50. Both transitivity effects---different term transitivity, and gwesp---are estimated to be positive and statistically significant in every year, and are even greater in magnitude than the reciprocity effect. The log-odds of a citation from case $i$ to case $j$ increases by more than 2 if there is at least one third case, $k$ that is cited by $i$, and cites to $j$. Results regarding the popularity effect (i.e., gwidegree) are generally supportive of the hypothesis of a popularity effect in Supreme Court citations, as the gwidegree estimate is negative in every term. There are some terms for which the result is not statistically significant at conventional levels. The popularity effect looks to have grown substantially stronger after the 2000 term. Contrary to expectations, we find that the hub effect is negative. That is, each additional citation sent by a case reduces the likelihood that the case is cited. For every citation sent, the log-odds of any citation to a case decreases by approximately 0.025---a result that is statistically significant in each term.

>>>>

We first discuss the dependence effect estimates in Figure \ref{SCC_results_1}. In most of the years that it is estimable, the reciprocity effect is statistically significantly positive, and substantial in magnitude. In a typical year, the presence of a citation from case $i$ to case $j$ increases the log odds of a citation from case $j$ to case $i$ by approximately 0.50. Both transitivity effects---different term transitivity and gwesp---are estimated to be positive and statistically significant in every year, and are even greater in magnitude than the reciprocity effect. The log odds of a citation from case $i$ to case $j$ increases by more than 2 if there is at least one third case, $k$ that is cited by $i$, and cites to $j$. Results regarding the popularity effect (i.e., gwidegree) are generally supportive of the hypothesis of a popularity effect in Supreme Court citations, as the gwidegree estimate is negative and statistically significant at conventional levels in nearly every term. This effect looks to have grown substantially after the 2000 term, which also coincides closely with the beginning of the Roberts court in 2006.

Contrary to expectations, we find that the hub effect is negative. That is, each additional citation sent by a case reduces the likelihood that the case is cited. For every citation sent, the log odds of any citation to a case decreases by an average of approximately 0.025---a result that is statistically significant in each term. This negative effect of receiver outdegree, notably, appears to be decreasing with time. Specifically, term to term variation aside, there is an upward trend toward zero in the parameter estimates. We believe this finding is particularly illustrative of why our approach of fitting separate models for each term is appropriate. As time passes, the number of nodes in the network (i.e., cases) grows. With the increasing number of citable cases and a relatively stable baseline rate of citation (i.e., the edges effect), there is an increase in outdegree mean and variation. This increasing difference between outdegree count by case makes it unlikely for the per unit change effect in outdegree to remain at a constant level. We have then an instance of network growth leading to changes in the model parameters even if the underlying network process remains unchanged.
